Extending the WebID Protocol with Access Delegation

### 2012-08-04 paper.tex

We can easily think of the origin server as not only able to respond to requests, but also as an agent able to make requests.
Indeed WebID authentication requires the server to make WebID profile requests to other servers in order to verify the identity of agents making a request to it.
The WebID specification describes this task as being accomplished by a separate agent, the WebID verifier - which could indeed be done by another service on the web (WebID proxy authentication servers).
But it can also be so closely tied to the web service that it would be natural to think of it as part of that same service.
The WebID profile furthermore could be served by the same agent as the one making the request, in which case we have a minimal case of a peer to peer communication.
Note that fetching a WebID profile for WebID authentication should be done anonymously, for fear of authentication deadlocks%
\footnote{For example one can imagine an agent $S$ with profile $P_s$ requesting a resource on server $R$ which requires authentication.
$S$ would send $R$ its certificate, thereby requiring $R$ to dereference $S$'s profile $P_s$ in order to verify the WebID.
If $P_s$ itself requires authentiction of $R$ and if $R$ sends a certificate containing a WebID with its profile $P_r$, and if $P_r$ itself requires authentication then it looks like we have a deadlock.}.

@@

\paragraph{Distinguish secretary agents from principal agents}
Delegation should not be handled by camouflaging an request of an agent in the sense that this request is indistinguishable from a principal agent request.
A secretary agent should use its own certificate (see efficency requirement) and it should be identified and described by its own WebID profile.
The motivation for this is variously.
(1) It allows resource guards to permit or deny requests based on this information.
(2) Applications which are secretary for many different principal agents do not need to switch their certificate between requests.
Also they do not need to handle hundreds of different keys.
(3) It allows for describing access control rules and the relation of a secretary agent to its principal agent.
Without a secretary WebID, this would not be possible.

@@

The simplest solution for any user agent $A$ wishing to act as the user $U$, is for it to create its own certificate $C_u$ with $U$'s WebID URI in the Subject Alternative Name position and add the public key to the user $U$'s profile $P_u$.
By doing this, the agent $A$ can then connect using its personal certificate $C_u$ and $U$'s profile, whenever it wants to work for / as the user $U$.
This solution has the advantage of reusing the existing WebID protocol in its current state, as well as that it transfers all powers of user $U$ to agent $A$. Therefore, this type of privilege should be limited to agents that are either extensions of $U$ or ones $U$ can or must fully trust anyway.

@@

In addition to the well known User Agents found in browsers on many different devices, there is one agent on which any service requiring WebID authentication relies on completely and this is the WebID profile's origin server~\cite{barth-a-2011--a}.
Any server requiring authentication of an agent using WebID must trust the server hosting the WebID profile, since the profile hosting server can change any information in the profile.
Any user profile can then be thought of as being a user according to the origin server i.e. a \textit{user@origin.org:port} - where \textit{origin.org} is the domain of the origin server, and \textit{port} is the port the server is running on.

When the user owns the origin server and there is a one to one relation between the two, as it is the case in a personal FreedomBox, then the origin server is just an extension of the user, and in that case trusting the server is the same as trusting the user.
%In that case the origin server can in fact just authenticate as the user whenever needed.

@@

Things get more interesting when an origin server agent $A$ is serving a number of different profiles $P_1$ \ldots $P_n$ each identifying respectively users $U_1$ \ldots $U_n$ and needs to act on behalf of each of these users.
This would be the situation for company servers, government agencies, educational institutions, charities, football clubs, etc.
This brings up two issues that need to be taken into account, an issue of data perspective and an issue of efficiency.

@@

In order to reduce the number of open connections to any server down to a minimum, it would be useful if the origin server could identify itself directly when making a request and specify on behalf of which of its users it was acting on per request so as to be able to make multiple requests on the same TLS connection.
One way would be for the origin server acting as client to simply use the same public key as the origin server acting as server - that is it could use the same public key as the one used by the TLS server.
It could even use the same certificate.
This would make clear to any server it was connecting to, that it could rely on this server as much as it could on the identity of any resource, and in particular any WebID profils served by that server.

Still, such an agent connecting to a remote service on behalf of one of its users would then need to identify which user it was acting on behalf of, or the remote service receiving a request would not know which access control rule to apply to the requesting origin server.
For example in our previous example in the Romeo and Juliet story we showed how the Capulet family Freedom Box only shows Juliet as knowing Romeo, when Romeo makes the request on the resource, but not when the request is made by Romeo's father, lord Montague.
That was an easy access control rule for the Capulet Freedbom Box to make when the request was using the WebID protocol simipliciter, since there it would know via the WebID authentication process exactly which identity the user was making the request as. 
But if it is the Origin Server connecting to the Capulet Freedom Box, identified as their Montague Origin Server, then what representation should the Capulet FB return? 
Should it return the reduced version available to Lord Montague?
Or should it return the more complete version available to Romeo and his close friends?

@@

The Capulet FB Guard having verified that this request does indeed come from the origin server of Romeo's profile
https://montague.net/romeo would then be able to accept that since the requesting server could in any case act as Romeo whenever it chose to do so, this protocol making exchange between the Montague and Capulet housholds more efficient without making it less secure, this be justification enough for it (the capulet freedom box) to  serve the representation that Romeo would have received had he connected to the server directly.  
We will argue further down that this is very similar to the way CORS deals with access control, namely by adding a header to a request to clarify what agent is making the request.

This On-Behalf-Of header makes it possible for one agent to open only one TLS connection to a remote server and be 
able to sepecify for each request on behalf of whome it is making it. The Montague Freedom Box could make a request to the capulet Freedom Box On-Behalf-Of Romeo, a few milliseconds later On-Behalf-Of Lord Montague, and so on... 
Again as explained in the previous section is is the responsibility of the Montague Freedom Box to keep these inforation seperate, and act with secrecy whenever it is acting on behalf of one or the other of the users. 
As such it's role is not unlike the role of Friar Laurence in the play - dedication to people's private matters requires tact and secrecy.

@@

If we now identify with a WebID the secretary that can act on behalf of a user, we can generalise the protocol somewhat.
That is, when an agent - such as the Montague Freedom Box Laurence with WebID \lstinline|<http://montague.net/fb#laurence>| - authenticates to a remote server it can use its own WebID. 
This would allow the Agent to have its own public key, and so for a compromise of the client public key not to affect the server public key.

But how would the remote server know that it can trust that secretary to be acting on behalf of a particular user?
It can no longer just compare the TLS keys of the requesting agent to see if it comes from the same Origin Server.
We need to make this relation explicity by use of a special RDF relation provisionally called \lstinline|cert:secretary|%
\footnote{A object property with a domain and range as \lstinline|foaf:Agent| which we will provisionally place
in the \lstinline|cert:| namespace, though it may be more appropriately placed in the \lstinline|auth:| namespace.}

@@

The remote server can then verify that the identified agent is the secretary of the agent he wishes to act on behalf of (as sepecified in the \lstinline|On-Behalf-Of| header, by dereferencing that users profile and verifying that the user specifies the \lstinline|cert:secretary| relation there, as it would if Romeo had the foaf profiles as shown in Listing \ref{listing:webid1}.

@@

%\autoref{fig:AuthSequence} depicts the process of a WebID authentication which is extended with the following components to allow access delegation:
The following enumeration describes each authentication step of \autoref{fig:AuthSequence} in detail but concentrates on the context of access delegation:
(1) The secretary agent opens a TLS connection with the server of the protected resource.
(2) Once TLS is set up, the HTTP request is sent to the server (e.g. a \verb!HTTP GET!) to consume or manipulate a resource.
        Here our first addition comes into play, the HTTP request header \lstinline|On-Behalf-Of| to claim an request as a secretary.
        We assume that the guard intercepts this HTTP request, and requests for a client authentication (using the TLS service)
(3) This authentication request is done using public key cryptography, sign a token with the guards private key and have the secretary send its certificate.
        This is defined in the TLS protocol which is specified in \cite{dierks-t-2012--a} and its successive RFCs.
        At the end of this step, the secretaries certificate is handled back to the guard.
(4) The guard ask the verification component to verify the secretary WebID which is named in the certificatei (5a).
This process is described in~\cite{story-h-2009--a} and is not touched by the delegation extension.
However, since the \lstinline|On-Behalf-Of| header was sent, the verifier component also fetches the WebID of the principal agent which is refered there.
(5b) The principal agents WebID is not verified in the same way as the secretary (e.g. no public key needed) but the secretary relation between secretary and principal agent is verified here.
For doing this, the verifier component simply needs to query an exact SPARQL ASK query for that triple.
(6) After this authentication process, the authorization process is executed.
This process is not topic of the paper and can involve a different resources.
An important issue here is, that the authorization has to be dann with the principal agent WebID and not with the secretary WebID.
(7) If the authorization process decides an access, the protected resource can be requested, otherwise a \lstinline|403 Forbidden| HTTP result is sent.

>>>>

We can easily think of the origin server as not only able to respond to requests, but also as an agent able to make requests.
Indeed WebID authentication requires the server to make WebID profile requests to other servers in order to verify the identity of agents making a request to it.
The WebID specification describes this task as being accomplished by a separate agent, the WebID verifier - which could indeed be done by another service on the web (WebID proxy authentication servers).
But it can also be so closely tied to the web service that it would be natural to think of it as part of that same service.
The WebID profile furthermore could be served by the same agent as the one making the request, in which case we have a minimal case of a peer to peer communication.
Note that fetching a WebID profile for WebID authentication should be done anonymously, for fear of authentication deadlocks%
\footnote{For example one can imagine an agent $S$ with profile $P_s$ requesting a resource on server $R$ which requires authentication.
$S$ would send $R$ its certificate, thereby requiring $R$ to dereference $S$'s profile $P_s$ in order to verify the WebID.
If $P_s$ itself requires authentication of $R$ and if $R$ sends a certificate containing a WebID with its profile $P_r$, and if $P_r$ itself requires authentication then it looks like we have a deadlock.}.

@@

\paragraph{Distinguish secretary agents from principal agents}
Delegation should not be handled by camouflaging an request of an agent in the sense that this request is indistinguishable from a principal agent request.
A secretary agent should use its own certificate (see efficiency requirement) and it should be identified and described by its own WebID profile.
The motivation for this is variously.
(1) It allows resource guards to permit or deny requests based on this information.
(2) Applications which are secretary for many different principal agents do not need to switch their certificate between requests.
Also they do not need to handle hundreds of different keys.
(3) It allows for describing access control rules and the relation of a secretary agent to its principal agent.
Without a secretary WebID, this would not be possible.

@@

The simplest solution for any user agent $A$ wishing to act as the user $U$, is for it to create its own certificate $C_u$ with $U$'s WebID URI in the Subject Alternative Name position and add the public key to the user $U$'s profile $P_u$.
By doing this, the agent $A$ can then connect using its personal certificate $C_u$ and $U$'s profile, whenever it wants to work for / as the user $U$.
This solution has the advantage of not making any modifications to the WebID protocol, as well as that it transfers all powers of user $U$ to agent $A$. However, this type of privilege should be limited to agents that are either extensions of $U$ or ones $U$ can or must fully trust anyway.

@@

%In addition to the well known User Agents found in browsers on many different devices, there is one agent on which any service requiring WebID authentication trusts implicitly and this is the WebID profile's origin server.
Any server requiring authentication of an agent using WebID must implicitly trust the origin server~\cite{barth-a-2011--a} hosting the WebID profile, since the profile hosting server can change any information in the profile.
%Any user profile can then be thought of as being a user according to the origin server, i.e. \textit{user@origin.org:port} - where \textit{origin.org} is the domain of the origin server, and \textit{port} is the port the server is running on.

When the user owns the origin server and there is a one-to-one relation between the two, as it is the case in a personal FreedomBox\footnote{http://www.freedomboxfoundation.org/}, then the origin server is just an extension of the user, and in that case trusting the server is the same as trusting the user.
%In that case the origin server can in fact just authenticate as the user whenever needed.

@@

Things become more interesting when an origin server agent $A$ is serving a number of different profiles $P_1$ \ldots $P_n$ each identifying respectively users $U_1$ \ldots $U_n$ and needs to act on behalf of each of these users.
This would be the situation for company servers, government agencies, educational institutions, charities, football clubs, etc.
This brings up two issues that need to be taken into account, an issue of data perspective and an issue of efficiency.

@@

In order to reduce the number of open connections to any server down to a minimum, it would be useful if the origin server could identify itself directly when making a request and specify on behalf of which of its users it was acting on per request so as to be able to make multiple requests on the same TLS connection.
One way would be for the origin server acting as client to simply use the same public key as the origin server acting as server - that is it could use the same public key as the one used by the TLS server.
It could even use the same certificate.
This would make clear to any server it was connecting to, that it could rely on this server as much as it could on the identity of any resource, and in particular any WebID profiles served by that server.

Still, such an agent connecting to a remote service on behalf of one of its users would then need to identify which user it was acting on behalf of, or the remote service receiving a request would not know which access control rule to apply to the requesting origin server.
For example in our previous example in the Romeo and Juliet story we showed how the Capulet family FreedomBox only shows Juliet as knowing Romeo, when Romeo makes the request on the resource, but not when the request is made by Romeo's father, lord Montague.
That was an easy access control rule for the Capulet FreedomBox to make when the request was using the WebID protocol, since there it would know via the WebID authentication process exactly which identity the user was making the request as. 
But if it is the Origin Server connecting to the Capulet (FreedomBox), identified as their Montague Origin Server, then what representation should the Capulet FreedomBox return? 
Should it return the reduced version available to Lord Montague?
Or should it return the more complete version available to Romeo and his close friends?

@@

The Capulet FreedomBox Guard having verified that this request does indeed come from the origin server of Romeo's profile
https://montague.net/romeo would then be able to accept that since the requesting server could in any case act as Romeo whenever it chose to do so, this protocol making exchange between the Montague and Capulet houses more efficient without making it less secure, this be justification enough for it (the Capulet FreedomBox) to  serve the representation that Romeo would have received had he connected to the server directly.  
In Section~\ref{sec:relatedwork} we will argue that this is very similar to the way CORS deals with access control, namely by adding a header to a request in order to clarify what agent is making the request.

This On-Behalf-Of header makes it possible for one agent to open only one TLS connection to a remote server and be 
able to specify for each request on behalf of whom it is making it. The Montague FreedomBox could make a request to the Capulet FreedomBox \textit{On-Behalf-Of} Romeo, a few milliseconds later On-Behalf-Of Lord Montague, and so on.
Again as explained in the previous section is is the responsibility of the Montague FreedomBox to keep these information separate, and act with secrecy whenever it is acting on behalf of one or the other of the users. 
As such it's role is not unlike the role of Friar Laurence in the play - dedication to people's private matters requires tact and secrecy.

@@

If we now identify with a WebID the secretary that can act on behalf of a user, we can generalise the protocol somewhat.
That is, when an agent - such as the Montague FreedomBox Laurence with WebID \lstinline|<http://montague.net/fb#laurence>| - authenticates to a remote server it can use its own WebID. 
This would allow the Agent to have its own public key, and so for a compromise of the client public key not to affect the server public key.

But how would the remote server know that it can trust that secretary to be acting on behalf of a particular user?
It can no longer just compare the TLS keys of the requesting agent to see if it comes from the same Origin Server.
We need to make this relation explicit by use of a special RDF relation provisionally called \lstinline|cert:secretary|%
\footnote{A object property with a domain and range as \lstinline|foaf:Agent| which we will provisionally place
in the \lstinline|cert:| namespace, though it may be more appropriately placed in the \lstinline|auth:| namespace.}

@@

The remote server can then verify that the identified agent is the secretary of the agent he wishes to act on behalf of (as specified in the \lstinline|On-Behalf-Of| header, by dereferencing that users profile and verifying that the user specifies the \lstinline|cert:secretary| relation there, as it would if Romeo had the FOAF profiles as shown in Listing \ref{listing:webid1}.

@@

%\autoref{fig:AuthSequence} depicts the process of a WebID authentication which is extended with the following components to allow access delegation:
The following enumeration describes each authentication step of \autoref{fig:AuthSequence} in detail but concentrates on the context of access delegation:
(1) The secretary agent opens a TLS connection with the server of the protected resource.
(2) Once TLS is set up, the HTTP request is sent to the server (e.g. a \verb!HTTP GET!) to consume or manipulate a resource.
        Here our first addition comes into play, the HTTP request header \lstinline|On-Behalf-Of| to claim an request as a secretary.
        We assume that the guard intercepts this HTTP request, and requests for a client authentication (using the TLS service)
(3) This authentication request is done using public key cryptography, sign a token with the guards private key and have the secretary send its certificate.
        This is defined in the TLS protocol which is specified in \cite{dierks-t-2012--a} and its successive RFCs.
        At the end of this step, the secretaries certificate is handled back to the guard.
(4) The guard ask the verification component to verify the secretary WebID which is named in the certificate (5a).
This process is described in~\cite{story-h-2009--a} and is not touched by the delegation extension.
However, since the \lstinline|On-Behalf-Of| header was sent, the verifier component also fetches the WebID of the principal agent which is referred there.
(5b) The principal agents WebID is not verified in the same way as the secretary (e.g. no public key needed) but the secretary relation between secretary and principal agent is verified here.
For doing this, the verifier component simply needs to query an exact SPARQL ASK query for that triple.
(6) After this authentication process, the authorization process is executed.
This process is not topic of the paper and can involve a different resources.
An important issue here is, that the authorization has to be done with the principal agent WebID and not with the secretary WebID.
(7) If the authorization process decides an access, the protected resource can be requested, otherwise a \lstinline|403 Forbidden| HTTP result is sent.
